EvoSwarm: A Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms

### 2020-05-15 draft.tex
Some comments here and there, check them out
Deletes some addressed comments and remarking others

  \title{EvoSwarm: A Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms}
  % maybe introduce "reproducible" here, somewhere, as well as "cloud native"
  % I like cloud-native more than event-based. 
  % How about that one?  - M

@@

The results are shown in Figure~\ref{fig:aws-results} and Table~\ref{tab:aws}.

As we can see, the local environment (see section \ref{setup}) achieved a higher
ratio of evaluations per second (in bold). Still, we must consider that there is a minimum
network overhead in a local workstation and a clock speed of 3.8 GHz. The first
option (labeled Single) is straightforward to deploy, needing only a single
{\tt docker-compose} file, and one command to execute. But with the limit of 4CPUs and
ten containers, it is the slower option. The second option (labeled Multi), is
faster, and not too far from the local execution, especially when solving the
$f_4$ in 20 dimensions (underline in table). With the capacity for increasing the number of worker
containers, it is reasonable to think that we can achieve a higher number of
evaluations with an appropriate configuration. Nevertheless, the goal of this
experiment was not to increase the number of evaluations but to test the
flexibility of a cloud-native application.

\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{aws}
  \caption{Boxplot of the number of evaluations per second, for 15 instances of function $f_4$,
   with three execution environments Local, Single task and Multi-task.
   }
  \label{fig:aws-results}
\end{figure}
%
\begin{table}[h!tbp]
  \small
  \caption{ Median number of the evaluations per second ratio, for 15 instances of function $f_4$,
  with three execution environments Local, Single task and Multi task}
  \label{tab:aws}
  \vspace{0.25cm}
  \centering

  \begin{tabular}{lll}
  \hline
  Dim & Env         & Median of ratio \\
  \hline
  10  & Local       & \textbf{47,166.29}       \\
      & Multi task  & 40,466.45    \\
      & Single task & 18,193.05      \\
  \hline
  20  & Local       & \textbf{36,101.86}   \\
      & Multi task  & {\ul 33,916.02}       \\
      & Single task & 16,546.95  \\
  \hline
  \end{tabular}
\end{table}

% You should clarify that, before, you were explaining how to set up an experiment. In this section, you are explaining the experiments that you are doing in this paper. But you should relate both sections to the general objectives of the paper. In the prior section, you want to prove that EvoSwarm can be used by the public to create reproductive research (and maybe you should add a small conclusion sub or sub-sub section), in the next you explain the experiment we are doing for this paper. It would be interesting to make references from the next section to the previous one, indicating how the general setup of EvoSwarm helps _us_ to create this experiment - JJ

@@

\item Does having a multi-population enabled platform, with heterogeneous populations and
the support for mixing search strategies,  increase the performance of the
search by needing fewer function evaluations than a homogeneous setting? % But in this case the answer is no, right? - JJ
                                                                         % Is yes, but is in the second experiment - M
\item Can EvoSwarm as a cloud-native, container-based application,  be deployed
locally or in the cloud by just specifying the configuration of container images
and resources to an orchestration service? % This does not fit in here... It's related to reproducibility, and as such belongs in the previous section. The subsections that answer this should also go to the previous section - JJ
\end{itemize}

@@

To validate these questions, we used benchmark functions from the Continuous
Noiseless BBOB testbed, which is part of the Comparing Continuous Optimizers
(COCO) framework \cite{hansen2016coco}. The testbed includes 24 real-parameter,
single-objective benchmark functions, and the capability to provide additional
instances of each function. Each instance of a function has a different optimal
value. The standard benchmark of the testbed utilizes 15 instances per function
over 2, 3, 5, 10, 20, and 40 dimensions. The maximum number of Function
Evaluations (FEs) changes according to the dimension (D), and it is determined
by the expression $10^5 \cdot D$. As an example, if we have $D = 2$, the
maximum number of FEs is $200,000$.
% How do you manage to stop exactly at that number of evaluations in a distributed system? - JJ
% In the controller we check the number of messages received, we assume that each message 
% needs a X amount of #FEs we stop at the maximum number of messages parameter.
% If the solution is found we stop sending messages and ignore messages from that
% experiment ID - M

@@

  \includegraphics[width=0.30\textwidth]{GAOnly_f005}&
  \includegraphics[width=0.30\textwidth]{PSOOnly_f005}&
  \includegraphics[width=0.30\textwidth]{GAPSO_f005}\\
  \end{tabular}
  \vspace{-3ex}
   \caption{
Scaling of expected runtime (RT defined as the number of function
evaluations to reach a target function value for the first time) for  
each problem dimension. Target values $\Delta f$ = 10, 1, $10^{-1}$, $10^{-2}$,$10^{-3}$,$10^{-5}$,$10^{-8}$
(the exponent and line colors are given in the legend of $f_1$ in the top row). 
Lines are the average RT (aRT); Cross ($+$): median RT of successful runs to reach
the most difficult target that was reached at least once (but not always);
Cross (\textcolor{red}{$\times$}): maximum number of f-evaluations in any trial. 
All values are divided by dimension and plotted as $log_{10}$ values versus dimension.
% More explanation for those not familiar with BBOB - JJ
% I think is more clear? - M
}
\label{fig:bbob}
\end{figure*}

>>>>

  \title{EvoSwarm: A Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms}
  % maybe introduce "reproducible" here, somewhere, as well as "cloud native"
  % I like cloud-native more than event-based. 
  % How about that one?  - M
  % better than the previous one.

@@

% Write some introduction to this subsubsection. 
The results are shown in Figure~\ref{fig:aws-results} and Table~\ref{tab:aws}.

As we can see, the local environment (see section \ref{setup}) achieved a higher
ratio of evaluations per second (in bold). Still, we must consider that there is a minimum
network overhead in a local workstation and a clock speed of 3.8 GHz. The first
option (labeled Single) is straightforward to deploy, needing only a single
{\tt docker-compose} file, and one command to execute. But with the limit of 4CPUs and
ten containers, it is the slowest option. The second option (labeled Multi), is
faster, and not too far from the local execution, especially when solving the
$f_4$ in 20 dimensions (underline in table). With the capacity for increasing the number of worker
containers, it is reasonable to think that we can achieve a higher number of
evaluations with an appropriate configuration. Nevertheless, the goal of this
experiment was not to increase the number of evaluations but to test the
flexibility of a cloud-native application.
% Mention costs - JJ

\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{aws}
  \caption{Boxplot of the number of evaluations per second, for 15 instances of function $f_4$,
   with three execution environments Local, Single task and Multi-task.
   }
  \label{fig:aws-results}
\end{figure}
%
\begin{table}[h!tbp]
  \small
  \caption{ Median number of the evaluations per second, for 15 instances of function $f_4$,
  with three execution environments Local, Single task and Multi task. The best result is shown in {\bf boldface}.}
  \label{tab:aws}
  \vspace{0.25cm}
  \centering

  \begin{tabular}{lll}
  \hline
  Dim & Env         & Median of ratio \\
  \hline
  10  & Local       & \textbf{47,166.29}       \\
      & Multi task  & 40,466.45    \\
      & Single task & 18,193.05      \\
  \hline
  20  & Local       & \textbf{36,101.86}   \\
      & Multi task  & {\ul 33,916.02}       \\ % why the underline?
      & Single task & 16,546.95  \\
  \hline
  \end{tabular}
\end{table}

% Emphasize these results for the purpose of the paper - JJ

% You should clarify that, before, you were explaining how to set up
% an experiment. In this section, you are explaining the experiments
% that you are doing in this paper. But you should relate both
% sections to the general objectives of the paper. In the prior
% section, you want to prove that EvoSwarm can be used by the public
% to create reproductive research (and maybe you should add a small
% conclusion sub or sub-sub section), in the next you explain the
% experiment we are doing for this paper. It would be interesting to
% make references from the next section to the previous one,
% indicating how the general setup of EvoSwarm helps _us_ to create
% this experiment - JJ

@@

\item Does having a multi-population enabled platform, with heterogeneous populations and
the support for mixing search strategies,  increase the performance of the
search by needing fewer function evaluations than a homogeneous setting? % But in this case the answer is no, right? - JJ
                                                                         % Is yes, but is in the second experiment - M
\item Can EvoSwarm as a cloud-native, container-based application,  be deployed
locally or in the cloud by just specifying the configuration of container images
and resources to an orchestration service? % This does not fit in
                                % here... It's related to
                                % reproducibility, and as such belongs
                                % in the previous section. The
                                % subsections that answer this should
                                % also go to the previous section - JJ
% Still need to check this out - JJ
\end{itemize}

@@

To validate these questions, we used benchmark functions from the Continuous
Noiseless BBOB testbed, which is part of the Comparing Continuous Optimizers
(COCO) framework \cite{hansen2016coco}. The testbed includes 24 real-parameter,
single-objective benchmark functions, and the capability to provide additional
instances of each function. Each instance of a function has a different optimal
value. The standard benchmark of the testbed utilizes 15 instances per function
over 2, 3, 5, 10, 20, and 40 dimensions. The maximum number of Function
Evaluations (FEs) changes according to the dimension (D), and it is determined
by the expression $10^5 \cdot D$. As an example, if we have $D = 2$, the
maximum number of FEs is $200,000$.

@@

  \includegraphics[width=0.30\textwidth]{GAOnly_f005}&
  \includegraphics[width=0.30\textwidth]{PSOOnly_f005}&
  \includegraphics[width=0.30\textwidth]{GAPSO_f005}\\
  \end{tabular}
  \vspace{-3ex}
   \caption{
Scaling of expected runtime (RT defined as the number of function
evaluations to reach a target function value for the first time) for  
each problem dimension. Target values $\Delta f$ = 10, 1, $10^{-1}$, $10^{-2}$,$10^{-3}$,$10^{-5}$,$10^{-8}$
(the exponent and line colors are given in the legend of $f_1$ in the top row). 
Lines are the average RT (aRT); Cross ($+$): median RT of successful runs to reach
the most difficult target that was reached at least once (but not always);
Cross (\textcolor{red}{$\times$}): maximum number of f-evaluations in any trial. 
All values are divided by dimension and plotted as $log_{10}$ values versus dimension.
% More explanation for those not familiar with BBOB - JJ
% I think is more clear? - M
% Almost everything, but the cream-colored rhombs in the bottom...
}
\label{fig:bbob}
\end{figure*}
